---
title: "Data Analysis Week 4"
subtitle: "Example Report"
author: "Robert Edwards"
output: 
  pdf_document:
        latex_engine: pdflatex
        number_sections: yes
fig_caption: yes
---

# Introduction {#sec:intro}
Section \ref{sec:intro}: This will link to he first section

## Subsection {#sec:sub}

# Code Chunks {#sec:code chunks}
Code "chunks" allow for R code to be embedded within a document. Not only can the code be easily included
within a document, the code can also be evaluated. Hence, you can produce an entire report based on an
analysis that is contained within a single file instead of having separate files containing your R code, plot
images and comments.
Using R code within an R Markdown document is done within the following environment:

```{r label, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
```
+ **echo**: include the R code within the code chunk in the document (TRUE/FALSE, default=TRUE);
+ **eval**: evaluate the R code within the code chunk (TRUE/FALSE, default=TRUE);
+ **warning**: suppress warnings from R (TRUE/FALSE, default=TRUE); and
3
+ **message**: suppress messages from R (TRUE/FALSE, default=TRUE).

Often the first use of a code chunk in a R Markdown file is to load packages so that functions and datasets
are available for analysis. For example, the following code chunk will load these packages:

+ <code>ggplot2</code> - to access the <code>tidyverse</code> visualization functions
+ <code>dplyr</code> - to access the <code>tidyverse</code> data manipulation functions
+ <code>skimr</code> - to access the <code>tidyverse</code> data summary functions
+ <code>moderndive</code> - to access the student feedback data <code>evals</code> that we looked at in Week 3
+ <code>datasets</code> - to access the <code>iris</code> data frame

```{r loadpackages, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(skimr)
library(moderndive)
library(datasets)
```

Then, let's say we wanted to select the <code>score</code> and <code>bty_avg</code> variables from the <code>evals</code> data set to be used
later, we can do that using the following code chunk:

```{r evals, echo = FALSE, eval = TRUE, warning = FALSE}
evals.scores <- evals %>%
select(score, bty_avg)
```

This will evaluate the R code and store the subsetted data set as the object evals.scores so that it can be
used later. If you want to embed the code within the Markdown document then you would simply set echo
= TRUE and it will include the R code as follows:

```{r evals2, echo=TRUE}
evals.scores <- evals %>%
dplyr::select(score, bty_avg)
```

## Inline Code{#sec:sub}
This section had errors

# Tables of Data{#sec:tables of data}
Let's say we wanted to create a table of the first 5 rows of the iris data from the datasets library. We can
create the table using the kable function as follows:

```{r table}
kable(head(iris, n = 5), caption = '\\label{tab:iris} The first 5 rows of the iris data.')
```

Notice that within the caption argument of the kable function there is <code>\\label{tab:iris}</code>. This is how
you label tables in order to refer to the within the text. For example,
<code>Table \ref{tab:iris}</code> displays the first 5 rows of the iris data...
will produce
<code>Table 1 displays the first 5 rows of the iris data...</code>
where, like for sections, the "1" will be a hyperlink directed to the table.

See also the Example Report Markdown file for an example of using the <code>kable_styling</code> function for changing
the size and positioning of a table. For example, its often useful to include <code>kable_styling(font_size =
10, latex_options = 'HOLD_position')</code> as the final term in the "pipe" to control the size of the font used
to construct the table and to ensure the table appears at that location in the document (otherwise it may be
placed at the bottom or top of the page).

## Table Summaries{#sec:sub}

Often we need to report summary statistics for different subsets in a set of data. For instance we can construct
a table of summaries for the sepal length of the different species in the iris data using:

```{r summaries, echo=TRUE}
library(kableExtra)
iris %>%
group_by(Species) %>%
summarise(n=n(),Mean=round(mean(Petal.Width),digits=1), St.Dev=round(sd(Petal.Width),digits=1),
Min=min(Petal.Width), Q1 = quantile(Petal.Width,0.25), Median=median(Petal.Width),
Q3 = quantile(Petal.Width,0.75), Max=max(Petal.Width)) %>%
kable(caption = '\\label{tab:summaries} Summary statistics on the sepal length by species of irises.') %>%
kable_styling(font_size = 10, latex_options = "hold_position")
```

## Tables of Model Estimates{#sec:sub}

Often we also need to report the results of fitting a model to our data. For instance if we modeled the sepal
length on the different species in the iris data by:

```{r fittedmodel}
model <- lm(Sepal.Length ~ Species, data = iris)
get_regression_table(model) %>%
dplyr::select(term,estimate) %>%
#Note that it seems necessary to include "dplyr::" here!!
kable(caption = '\\label{tab:reg} Estimates of the parameters from the fitted linear regression model.') %>%
kable_styling(latex_options = 'HOLD_position')
```

## Tables By Hand{#sec:sub}

Tables can also be produced "by hand"" in Markdown. For example, the table above corresponding to the
first 5 rows of the iris data can be produced by hand by typing the following text (without any other text)
into a .Rmd file:

Sepal Length  | Sepal Width
:------------:|:------------:
1             |2
3             |4
5             |6
7             |8
9             |10

The vertical separators | are used between columns, while --- is placed below table/column headings.
Alignment of the columns is done using colons, that is, for left alignment put :---, for right alignment put
---:, and for centred alignment put :---:. A limitation of creating tables by hand is that you cannot give
them a label to refer to them later within the text (which will also need to be done by hand).


## Figures{#sec:figures}

Including plots within an R Markdown document is straightforward. The R code for the plot is simply
included within a code chunk including additional arguments for plot size and positioning. For example, this
code chunk in the .Rmd file will produce the scatterplot of teaching and beauty scores below:

```{r scatplot, echo = TRUE, eval = TRUE, out.width = '70%', fig.align = "center", warning = FALSE, fig.cap = "\\label{fig:scat} Relationship between teaching and beauty scores. The best-fitting line has been superimposed."}
ggplot(evals.scores, aes(x = bty_avg, y = score)) +
geom_point() +
labs(x = "Beauty Score", y = "Teaching Score") +
geom_smooth(method = "lm", se = FALSE)
```

Here, we have <code>echo = FALSE</code> and <code>eval = TRUE</code> as we only want to put the plot into the document and not the
accompanying R code. Labelling is done the same way as with tables, that is, the label <code>\\label{fig:scat}</code>
is placed within the figure caption, which is given by the argument fig.cap. For size and positioning of the
figure we can include:
+ <code>out.width</code>: a percentage of the actual size of the produced plot;
+ <code>fig.width</code>: an integer value denoting the width of the figure;
+ <code>fig.align</code>: the alignment of the figure within the body of the document; and
+ <code>fig.pos</code>: can be used to force the positioning of the figure within the document.
For examples of using each of the above arguments see the Example Report Markdown file <code>ExampleReport.Rmd</code>.

# Mathematics{#sec:math}











