---
title: "Predicting Medical Insurance Charges from BMI and Smoking Status"
author: "DA Group 6"
output: 
  pdf_document:
          latex_engine: pdflatex
          number_sections: yes
fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA)
options(scipen = 999)
```

```{r libraries}
library(ggplot2)
library(dplyr)
library(moderndive)
library(skimr)
library(tidyr)
library(kableExtra)
library(gridExtra)
library(janitor)
library(infer)
library(broom)
```


```{r data, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
insurance <- read.csv("insurance.csv")
insur<- insurance%>%
 filter(region == "southeast")%>%
 select(charges, age, bmi, smoker)
```

# Introduction {#sec:Intro}

In today's world, medical care is undoubtedly an expensive affair.  Consequently, numerous individuals enroll in a health insurance policy, where they agree to pay a sum of money (known as premiums) to a particular health insurance company, on a monthly or yearly basis.  In return, the company will guarantee to reimburse a proportion of the medical costs in case the insured is injured or sick and needs medical treatment. 

The amount of premium which an individual pays, depends on a number of factors.  In particular, the higher the risk of having health problems, the higher the premium would be.  For instance, overweight or obese persons are more likely to develop heart disease when compared with people of normal weight.  Moreover, smoking is linked with an increased risk of lung cancer when compared to individuals who do not smoke.  Apart from these, there is a tendency for health insurance rates to escalate with increasing age, since older people are more prone to health problems.  In these situations, one might expect these individuals to have a higher risk of large medical expenses.

The main goal of this report is to predict the total yearly medical costs (`charges`) billed by health insurance from the age of the individual (`age`), the corresponding body mass index (`bmi`) and from whether the insured smokes tobacco or not (`smoker`).  The `bmi` is a measure of body fat, which is defined as the body weight (in kg) divided by the square of the body height (in $m^2$).  The data analysed here consists of a random sample of `r nrow(insur)` individuals from 1338 observations, which contain hypothetical data on medical expenses for patients in the United States.  The simulated data was based on the demographic statistics obtained from the US Census Bureau, and thus, it resembles real world data.  

Section \ref{sec:EDA} consists of an exploratory data analysis to gain a better understanding of the distribution of the features used under this study.  Section \ref {sec:FDA} is divided into two parts.  The first subsection provides the results which are obtained after fitting a multiple linear regression model to the data.  Moreover, the model assumptions are also checked here.  Then, in the second subsection, confidence intervals are calculated in order to obtain a range of plausible values for the average medical care expenses.  Finally, Section \ref{sec:Conc} sums up the results obtained after conducting this analysis.

```{r summary1, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

insur %>%
  group_by(smoker) %>%
  summarise(n()) %>%
  kable(col.names = c("Smoker", "n"),
        caption = '\\label{tab:sum1} Number of smokers and non-smokers.', booktabs = TRUE, format = "latex") %>%
  kable_styling(font_size = 10, latex_options = "hold_position")
```

# Exploratory Data Analysis {#sec:EDA}
Table \ref{tab:summary1} shows that there are `r sum(insur$smoker=="yes")` of smokers involves in the study which is much smaller than non smoker participant  which equal to `r sum(insur$smoker=="no")`. On the other hand, Table \ref{tab:summary2} contains summary statistics on the age, bmi and charges. Start by looking at the mean age of individuals are `r round(mean(insur$age), 2)`, additionally,  50% of the body mass index (bmi) lies between `r round(quantile(insur$bmi, 0.25), 2)` and `r round(quantile(insur$bmi, 0.75), 2)`. If we look at medical costs (charges) we see that the middle 50% lies between `r round(quantile(insur$charges, 0.25), 2)` and `r round(quantile(insur$charges, 0.75), 2)` dollars, with an average of `r round(mean(insur$charges), 2)` dollars. The standard deviation are `r round(sd(insur$age), 2)`, `r round(sd(insur$bmi), 2)` and `r round(sd(insur$charges), 2)`, which is the variation of observations about the mean for age, bmi and charges respecevily. 

```{r summary2, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

insur$age <- as.numeric(insur$age)

skim_with(numeric = list(hist = NULL, missing = NULL, complete = NULL))
insur %>%
  skim_to_list() %>%
  .$numeric %>%
  kable(col.names = c("Variable", "n", "Mean", "SD", "Minimum", "1st quartile", "Median", "3rd quartile", "Maximum"),
        caption = '\\label{tab:sum2} Summary statistics on Insurance charges and bmi.', booktabs = TRUE, format = "latex") %>%
  kable_styling(font_size = 10, latex_options = "hold_position")
```
```{r corr.coef, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
Cor.1 <- insur %>%
  get_correlation(formula = charges ~ age)

Cor.2 <- insur %>%
  get_correlation(formula = charges ~ bmi)

Cor.3 <- insur %>%
  get_correlation(formula = bmi ~ age)
```
In order to check the collinearity between the continous variables in the study, the pair scatterplot plotted in Figure \ref {fig:cor}. which shows that there no relationship between any two continous variables. For confirmation the correlation coefficent calculated between variables shows the correlation  between charges and age is (`r round(Cor.1, 3)`), also the correlation between charges and bmi is (`r round(Cor.2, 3)`) . And the association between the two continous explanatory variables bmi and age is (`r round(Cor.2, 3)`). Moreover, from the correlation coefficient we can consider that we have very weak positiove relationship between continous variables, which indicate that there is no collinearity in the study.

```{r correlation, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, fig.align="center", fig.pos="H",fig.cap = "\\label{fig:cor} scatterplot pair plot between charges, age and bmi"}

d<- insur[,1:3]
pairs(d, col="aquamarine3")


```

In Figure \ref {fig:scat} a linear regression model was fitted in order to assess the association between response variable (insurance charges) and explanatory variables (age, bmi and smoking status).
Start with left hand plot, which shows the relationship between charges and bmi by the smoking status of individuals. It appear from the plot that as the body mass index(bmi) increase, smokers insurance charge increase dramatically. On the other hand, non-smoker remain stable as their body mass index increase. by looking at the right hand plot, which explain the relationship between charges and age by the smoking status. We can see that as people muture they pay more health insurance charge. However, non-smoker pay less than 40000 dollar where smokers pay up to 60000 dollar on their health insurance. To sum up, it is evident from both scatterplots that smoking people pay lage amount on their health insurance comparing with those who dont smoke.
A linear regression model will be fitted to assess the relationship between charges and age, bmi and smoking status significantly.

```{r scatplot, fig.width=13, fig.pos = "H", fig.align = "center", fig.cap = "\\label{fig:scat} scatterplots of insurance charges and bmi by smoking status (left) and insurance charges and age by smoking status (right)"}
p1 <- ggplot(insur, aes(x = bmi, y = charges, color = smoker)) +
  geom_point() +
  labs(x = "bmi", y = "Insurance charges (in $)", color = "smoker") +
  geom_smooth(method = "lm", se = FALSE)

p2 <- ggplot(insur, aes(x = age, y = charges, color = smoker)) +
  geom_point() +
  labs(x = "age", y = "Insurance charges (in $)", color = "smoker") +
  geom_smooth(method = "lm", se = FALSE)

grid.arrange(p1, p2, ncol =2)
```

# Formal Data Analysis {#sec:FDA}

The parallel regression lines model is given as follows:

$$y_i = \alpha + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{3i} + \epsilon_i$$
$$y_i = \alpha + \beta_{\mbox{bmi}} \cdot \mbox{bmi}_{i} +\beta_{\mbox{age}} \cdot \mbox{age}_{i} + \beta_{\mbox{Smoker}} \cdot \mathbb{I}_{\mbox{Smoker}}(i) + \epsilon_i $$
where

- $\alpha$ is the intercept of the regression line for non-smokers;
- $\beta_{\mbox{bmi}}$ is the slope of the regression line for both smokers and non-smokers;
- $\mbox{bmi}_{i}$ is the bmi of the $i^{th}$ observation
- $\beta_{\mbox{age}}$ is the slope of the regression line for both smokers and non-smokers;
- $\mbox{age}_{i}$ is the age of the $i^{th}$ observation
- $\beta_{\mbox{Smoker}}$ is the additional term added to $\alpha$ to get the intercept of the regression line for smokers; and
- $\mathbb{I}_{\mbox{Smoker}}(i)$ is an indicator function such that

$$\mathbb{I}_{\mbox{Smoker}}(i)=\left\{
\begin{array}{ll}
1 ~~~ \mbox{if the} ~ i \mbox{th observation smokes}, \\
0 ~~~ \mbox{Otherwise}.\\
\end{array}
\right.$$

The parameter estimates obtained in the parallel model are shown in Table \ref{tab:reg}:

```{r fittedmodel, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

par.model <- lm(charges ~  bmi + age+ smoker, data = insur)

coeff <- par.model %>%
  coef() %>%
  as.numeric()

get_regression_table(par.model) %>%
  kable(caption = '\\label{tab:reg} Estimated Coefficients from the fitted model charges = bmi +age+ smoker', booktabs = TRUE, format = "latex") %>%
  kable_styling(font_size = 10, latex_options = 'HOLD_position')
reg.tab1<- get_regression_table(par.model)
reg1<- as.data.frame(reg.tab1)
```

From Table \ref {tab:reg} we can see that the  p-value<0.001, which indicate that all parametrs in the simple linear equation is significant. Also the confidence interval of $\beta_{\mbox{bmi}}$is  (`r reg1[2,6]`, `r reg1[2,7]`), the confidence interval of $\beta_{\mbox{age}}$is  (`r reg1[3,6]`, `r reg1[3,7]`) and the confidnce interval of  $\beta_{\mbox{smoke}}$is  (`r reg1[4,6]`, `r reg1[4,7]`). which dont contain zero.
Hence, the regression line for non-smokers is given by:

$$\widehat{\mbox{Charges}} = `r round(coeff[1],3)` + `r round(coeff[2],3)`\cdot{\mbox{bmi}} + `r round(coeff[3],3)`\cdot{\mbox{age}} $$ 

while the regression line for smokers is given by:

$$\widehat{\mbox{Charges}} = `r round(coeff[1],3)` + `r round(coeff[2],3)`\cdot{\mbox{bmi}} +`r round(coeff[3],3)`\cdot{\mbox{age}} + `r round(coeff[4],3)` = `r round(coeff[1]+coeff[4],3)` + `r round(coeff[2],3)`\cdot{\mbox{bmi}}+ `r round(coeff[3],3)`\cdot{\mbox{age}}$$ 


The interaction regression lines model is given as follows:

```{r fittedmodel2, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
int.model <- lm(charges ~ age + bmi + age*smoker + bmi*smoker, data = insur)

coeff <- int.model %>%
  coef() %>%
  as.numeric()

get_regression_table(int.model) %>%
  kable(caption = '\\label{tab:reg2} Estimated Coefficients from the interaction fitted model', booktabs = TRUE, format = "latex") %>%
  kable_styling(font_size = 10, latex_options = 'HOLD_position')

reg.tab2<- get_regression_table(int.model)
reg2<- as.data.frame(reg.tab2)
```

From Table \ref {tab:reg2} we conclude that there is insufficent evedent that $\beta_{\mbox{bmi}}$ and  $\beta_{\mbox{age, smokeryes}}$ differ from zero as their confedice interval contain zero,(`r reg2[3,6]`, `r reg2[3,7]`) and (`r reg2[5,6]`, `r reg2[5,7]`) confidence interval of $\beta_{\mbox{bmi}}$ and  $\beta_{\mbox{age, smokeryes}}$ respectivly.
Now,the model selection will be conducted to find the best fitted line.

```{r model.sel, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

model.comp.values.full <- glance(lm(charges ~ age * bmi * smoker, data = insur))

model.comp.values.m1 <- glance(lm(charges ~ age + bmi + smoker + age*bmi + age*smoker + bmi*smoker, data = insur))

model.comp.values.m2 <- glance(lm(charges ~ age + bmi + smoker + age*bmi + bmi*smoker, data = insur))

model.comp.values.m3 <- glance(lm(charges ~ age + bmi + smoker + bmi*smoker, data = insur))

model.comp.values.m4 <- glance(lm(charges ~ age + smoker, data = insur))

Models <- c('Full Model', 'Model 1', 'Model 2', 'Model 3', 'Model 4')

bind_rows(model.comp.values.full, model.comp.values.m1, model.comp.values.m2, model.comp.values.m3, model.comp.values.m4, .id = "Model") %>%
  select(Model, adj.r.squared, AIC, BIC) %>%
  mutate(Model = Models) %>%
   kable(caption = '\\label{tab:bestmodel} model selection', booktabs = TRUE, format = "latex") %>%
  kable_styling(font_size = 10, latex_options = 'HOLD_position')
  
```

From Table \ref {tab:bestmodel} the model the model with largest adjusted R squared and lowest AIC and BIC is the best fitted model and we can conclude that the best fitted regression model is 
$$y_i = \alpha + \beta_{\mbox{bmi}} \cdot \mbox{bmi}_{i} +\beta_{\mbox{age}} \cdot \mbox{age}_{i} + \beta_{\mbox{Smoker}} \cdot \mathbb{I}_{\mbox{Smoker}}(i) + \beta_{\mbox{bmi, smoker}} \cdot \mathbb{I}_{\mbox{bmi, smoker}}(i) +\epsilon_i $$
Where $$ \beta_{\mbox{bmi, smoker}} \cdot \mathbb{I}_{\mbox{bmi, smoker}}(i)$$ correspond to the interaction term.

```{r bestfittedmodel, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
modelb <- lm(charges ~ age + bmi + bmi*smoker, data = insur)

get_regression_table(modelb) %>%
  kable(caption = '\\label{tab:best} Estimated Coefficients from the best fitted model', booktabs = TRUE, format = "latex") %>%
  kable_styling(font_size = 10, latex_options = 'HOLD_position')

reg.tab<- get_regression_table(modelb)
reg<- as.data.frame(reg.tab)
```

From the summary statsitic of the best fitted model we can consider that we dont have enough evedince to reject that $\beta_{\mbox{bmi}}$ is differ from zero as the confidnce inetrval of $\beta_{\mbox{bmi}}$ is (`r reg[3,6]`, `r reg[3,7]`) which contain zero.  Hence, we need to examine the assumptions for the best fitted model.


```{r residplot11, echo=FALSE, out.width = '75%', fig.align = "center", fig.cap = "\\label{fig:resid1} Residuals vs the explanatory variable bmi by Smokers.", message = FALSE, fig.pos="H"}

modelb <- lm(charges ~ age + bmi + bmi*smoker, data = insur)
regression.points1 <- get_regression_points(modelb)

w1<-ggplot(regression.points1, aes(x = bmi, y = residual)) +
  geom_jitter(color="aquamarine3") +
  labs(x = "bmi", y = "Residual") +
  geom_hline(yintercept = 0, col = "blue", size = 1) +
  facet_wrap(~ smoker)
w2<-ggplot(regression.points1, aes(x = age, y = residual)) +
  geom_jitter(color="coral1") +
  labs(x = "age", y = "Residual") +
  geom_hline(yintercept = 0, col = "blue", size = 1) +
  facet_wrap(~ smoker)
grid.arrange(w1, w2, ncol =2)

```
By looking at the scatterplot \ref {fig:resid1} of residuals against the explanatory variables by smoking statu, we can see that the assumptions of residuals having mean equal to zero and consitant variability across all levels of the explanatory variable holds.

```{r residplot22, echo=FALSE, out.width = '75%', fig.align = "center", fig.cap = "\\label{fig:resid22} Residuals vs the fitted values by Smoker.", message = FALSE, fig.pos="H"}
ggplot(regression.points1, aes(x = charges_hat, y = residual)) + 
  geom_point(color="coral1") +
  labs(x = "Fitted values", y = "Residual") +
  geom_hline(yintercept = 0, col = "blue", size = 1) +
  facet_wrap(~ smoker)
```
Figure \ref {fig:resid22} shows scatterplot of residuals against fitted values. Which we can see that there is no obvious pattern in the residuals and that confirm the zero mean and consitant variance condistion. Hence the random error component of the regression model staisfies.

```{r residplot33, echo=FALSE, out.width= '75%', fig.align = "center", fig.cap = "\\label{fig:resid33} Histogram of residuals.", message = FALSE, fig.pos="H"}
ggplot(regression.points1, aes(x = residual)) + 
  geom_histogram(color = "white", fill="aquamarine3") +
  labs(x = "Residual") +
  facet_wrap(~ smoker)
```

The histogram of the residuals shown in \ref {fig:resid33} indicate that the residuals appears to be relatively bell-shaped and centred around zero.Therefore, the assumptions of normallity distributed errors appears to be hold for the best fitted model. 


# Conclusions {#sec:Conc}