---
title: "Practice Class Test2 Template"
author: "2416963E"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
library(ggplot2)
library(GGally)
library(gridExtra)
library(dplyr)
library(moderndive)
library(ISLR)
library(skimr)
library(plotly)
library(tidyr)
library(datasets)
library(knitr)
library(janitor)
library(infer)
library(broom)
library(sjPlot)
library(stats)
library(readr)
```

## Question 1
Large samples of iron ore were mined from quarries "x" and "y", and each of the two samples were broken down into 10 smaller sub-samples for analysis. Each of these 20 sub-samples were sent to a chemical laboratory, and the percentage of iron in each sub-sample was measured. These data are stored in `practice1.csv`. Using bootstrapping, is there evidence in these data that the population mean iron percentage in each quarry is 35%, and are the population mean percentages different between the two quarries?

```{r, warning=FALSE, message=FALSE, echo=FALSE}
q1data <- read_csv("practice1.csv")
```


## Question 2

(a) Using the data contained in `practice2.csv` build a regression model that adequately describes the `response` in terms of the potential explantory variables `X1`, `X2`, `X3` and `X4`. Your chosen model will therefore be the one that you believe best represents the `response`. Use only the theoretical confidence intervals generated under standard assumptions (which you should check) to identify the correct model.  Note you **do not** need to consider interaction terms.

```{r, message=FALSE, echo=FALSE}
q2data <- read_csv("practice2.csv")
q2data$X2 <- as.factor(q2data$X2)
```

(b)  Construct a table of **all** the possible linear models (without interactions or transformations) that could be fitted to the `response` variable in `practice2.csv`.  In the table include the $R^2_{adj}$ and $AIC$ values for model comparisons. Do these measures lead you to the same conclusion about the model that best represents the data as in part (a)?  Note: you are **not** required to check assumptions for each of these models in this task.

