---
title: "Practice Class Test2 Template"
author: "2416963E"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
library(ggplot2)
library(GGally)
library(gridExtra)
library(dplyr)
library(moderndive)
library(ISLR)
library(skimr)
library(plotly)
library(tidyr)
library(datasets)
library(knitr)
library(janitor)
library(infer)
library(broom)
library(sjPlot)
library(stats)
library(readr)
library(kableExtra)
```

## Question 1
Large samples of iron ore were mined from quarries "x" and "y", and each of the two samples were broken down into 10 smaller sub-samples for analysis. Each of these 20 sub-samples were sent to a chemical laboratory, and the percentage of iron in each sub-sample was measured. These data are stored in `practice1.csv`. Using bootstrapping, is there evidence in these data that the population mean iron percentage in each quarry is 35%, and are the population mean percentages different between the two quarries?

```{r, eval = TRUE,  echo=TRUE, warning=FALSE, message=FALSE}
q1data <- read_csv("practice1.csv")
data1 <- gather(data = q1data,
                key = ore,   ## select column to collapse
                value = percent, ## column name for values
                1:2)
```


Part (a) asks if there is evidence in these data that the population mean iron percentage in each quarry, $\mu_x$ and $\mu_y$, is 35% using bootstramp sampling.  We can test each of these with a 95% confidence interval on the bootstrap sample means, $\bar{x}_x$ and $\bar{x}_y$ respectively.   

```{r echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
samples_x <- data1 %>% 
  filter(ore == "x") %>%
  specify(formula = percent ~ NULL) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "mean")

samples_y <- data1 %>% 
  filter(ore == "y") %>%
  specify(formula = percent ~ NULL) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "mean")

x_bar <- data1 %>%
  filter(ore == "x") %>%
  summarize(stat = mean(percent))
y_bar <- data1 %>%
  filter(ore == "y") %>%
  summarize(stat = mean(percent))

px <- samples_x %>%
  visualize(obs_stat = x_bar)
py <- samples_y %>%
  visualize(obs_stat = y_bar)

ci_x <- samples_x %>%
  get_ci(level = 0.95, type = "percentile")
ci_y <- samples_y %>%
  get_ci(level = 0.95, type = "percentile")

#px <- samples_x %>%
#  visualize(endpoints = ci_x, direction = "between")
#py <- samples_y %>%
#  visualize(endpoints = ci_y, direction = "between")

# Bootstrap sample is close to symmetric and bell-shaped (Normal) so we can use the Standard Error method
ci_x <- samples_x %>%
  get_ci(level = 0.95, type = "se", point_estimate = x_bar)
ci_y <- samples_y %>%
  get_ci(level = 0.95, type = "se", point_estimate = y_bar)

px <- samples_x %>%
  visualize(endpoints = ci_x, direction = "between")
py <- samples_y %>%
  visualize(endpoints = ci_y, direction = "between")

```

Since the distribution of bootstrap samples of x and y are both close to symmetric and bell-shaped a 95% confidence interval was created using the standard error method. The 95% confidence interval for x is [`r round(ci_x[1], 2)`, `r round(ci_x[2], 2)`].  The confidence interval does not include 0.35 (35%) and we conclude that there is not evidence in these data that the population mean iron percentage in quarry x is 35%. For quarry y the 95% confidence interval is [`r round(ci_y[1], 2)`, `r round(ci_y[2], 2)`].  The confidence interval includes 0.35 and we conclude there is evidence in these data that the population mean iron percentage in quarry y is 35%.  

Part (b) asks if there is evidence in these data that the population mean is difference between the two quarries.  

```{r echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
set.seed(222)

samples_xy <- data1 %>% 
  specify(formula = percent ~ ore) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "diff in means", order = c("x", "y"))


samples_xy %>%
  visualize(obs_stat = x_bar - y_bar)

ci_xy <- samples_xy %>%
  get_ci(level = 0.95, type = "percentile")
```

The bootstrap samples are created by subtracting y from x.  The 95% confidence interval is made using the percentiles method and is found to be  [`r round(ci_xy[1], 2)`, `r round(ci_xy[2], 2)`], which contains zero.  Since the confidence interval contains zero we conclude that there is no evidence that the two quarries are different.  


## Question 2

(a) Using the data contained in `practice2.csv` build a regression model that adequately describes the `response` in terms of the potential explantory variables `X1`, `X2`, `X3` and `X4`. Your chosen model will therefore be the one that you believe best represents the `response`. Use only the theoretical confidence intervals generated under standard assumptions (which you should check) to identify the correct model.  Note you **do not** need to consider interaction terms.

```{r, message=FALSE, echo=FALSE}
data2 <- read_csv("practice2.csv")
data2$X2 <- as.factor(data2$X2)
```

First, a pairs plot, Figure \ref{fig:pairs1}, is made to explore the data visually and see if there are any patterns or correlations in the data. The data consists of 

```{r pairs2, echo = FALSE, eval = FALSE, fig.width=7, fig.height=3,  fig.pos="H", fig.align = "center", warning = FALSE, fig.cap = "\\label{fig:pairs1}Pairs plot for question 2."}
ggpairs(
  data2,
  mapping = ggplot2::aes(color = X2),
  upper = list(continuous = wrap("density", alpha = 0.5), combo = "box"),
  lower = list(continuous = wrap("points", alpha = 0.4), combo = wrap("dot", alpha = 0.4)),
  diag = list(continuous = wrap("densityDiag"))
  )
```

```{r, message=FALSE, echo=FALSE}
model1 <- lm(response ~ X1 + X2 + X3 + X4, data = data2)

get_regression_table(model1) %>%
  kable(digits = 3, caption = '\\label{tab:fullmodel} Parameter estimates obtained from the model response ~ X1 + X2 + X3 + X4', booktabs = TRUE, format = "latex") %>%
  kable_styling(font_size = 10, latex_options = "hold_position")

reg.tab1<- get_regression_table(model1)
reg1<- as.data.frame(reg.tab1)
```

(b)  Construct a table of **all** the possible linear models (without interactions or transformations) that could be fitted to the `response` variable in `practice2.csv`.  In the table include the $R^2_{adj}$ and $AIC$ values for model comparisons. Do these measures lead you to the same conclusion about the model that best represents the data as in part (a)?  Note: you are **not** required to check assumptions for each of these models in this task.

```{r modelselection, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}

model.comp.full <- glance(lm(response ~ X1 + X2 + X3 + X4, data = data2))
model.comp.123 <- glance(lm(response ~ X1 + X2 + X3, data = data2))
model.comp.12 <- glance(lm(response ~ X1 + X2, data = data2))
model.comp.13 <- glance(lm(response ~ X1 + X3, data = data2))
model.comp.14 <- glance(lm(response ~ X1 + X4, data = data2))
model.comp.23 <- glance(lm(response ~ X2 + X3, data = data2))
model.comp.24 <- glance(lm(response ~ X2 + X4, data = data2))
model.comp.34 <- glance(lm(response ~ X3 + X4, data = data2))
model.comp.4 <- glance(lm(response ~ X4, data = data2))
model.comp.3 <- glance(lm(response ~ X3, data = data2))
model.comp.2 <- glance(lm(response ~ X2, data = data2))
model.comp.1 <- glance(lm(response ~ X1))

Models <- c('Full Model', 'Model 123', 'Model 234', 'Model 134', 'Model 234', 'Model 12', 'Model 13', 'Model 14', 'Model 23', 'Model 24', 'Model 34', 'Model 1', 'Model 2', 'Model 3', 'Model 4')

bind_rows(model.comp.values.full, model.comp.values.m1, model.comp.values.m2, model.comp.values.m3, model.comp.values.m4, .id = "Model") %>%
  select(Model, adj.r.squared, AIC, BIC) %>%
  mutate(Model = Models) %>%
   kable(caption = '\\label{tab:select} Model selection', booktabs = TRUE, format = "latex") %>%
  kable_styling(font_size = 10, latex_options = 'HOLD_position')
```
